Understanding the Impact of Marital Status, Income, and Age on Individuals' Poverty Status

### 2024-03-29 paper/paper.qmd
change to marital status

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Mary Cheng
thanks: "Code and data are available at: https://github.com/marycx/us_poverty_analysis_2019.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}

```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

@@

# Convert variables to factors
analysis_data$mortgage <- factor(analysis_data$mortgage)
analysis_data$income <- factor(analysis_data$income)
analysis_data$age <- factor(analysis_data$age)

@@

posterior_vs_prior(poverty_prediction_model) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and mortgage"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of owner without mortgage", "Trace plot of renter"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "(Intercept)")
plot(poverty_prediction_model, "trace", "mortgageOwner without mortgage")
plot(poverty_prediction_model, "trace", "mortgageRenter")
```

@@

plot(poverty_prediction_model, "rhat")
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"

>>>>

---
title: "Understanding the Impact of Marital Status, Income, and Age on Individuals' Poverty Status"
subtitle: "An Analysis of the 2019 Revised Supplemental Poverty Measure Dataset"
author: 
  - Mary Cheng
thanks: "Code and data are available at: https://github.com/marycx/us_poverty_analysis_2019.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

## Data Source

## Features

# read in raw data
raw_poverty_data <-  read_csv(here::here("data/raw_data/raw_poverty_data.csv"))
```

```{r}
#| label: tbl-raw
#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

raw_poverty_data[1:5, 1:4] |>
  kable()
```

## Data Measurement

## Methodology

# read in cleaned data
poverty2019 <-  read_parquet(here::here("data/analysis_data/cleaned_poverty_data.parquet"))

poverty2019 <-
  poverty2019 |>
  mutate(
    poverty_status = as_factor(poverty_status),
    marital_status= factor(
      marital_status,
      levels = c(
      "Married - civilian spouse present",
      "Married - armed forces spouse present",
      "Married - spouse absent (excluding separated)",
      "Widowed",
      "Divorced",
      "Separated",
      "Never Married")
    ),
    income = factor(
      income,
      levels = c(
      "below 10k",
      "10k-50k",
      "50k-100k",
      "100k-150k",
      "150k-200k",
      "200k-250k",
      "above 250k")
    ),
    age = factor(
      age,
      levels = c(
      "25-34",
      "35-44",
      "45-54",
      "55-64",
      "above 65")
    )
  ) |>
  select(poverty_status, marital_status, income, age)
```

```{r}
#| label: tbl-clean-data
#| tbl-cap: "Preview of the cleaned 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

poverty2019[5:9, 1:4] |>
  kable()
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2019 US Supplemental Poverty Measure dataset"
#| message: false
#| echo: false

poverty2019 |>
  summary() |>
  kable(align = "c") %>%
  kable_styling(font_size = 8)
```

## Data Visualization
```{r}
#| label: fig-mortgage
#| fig-cap: The distribution of poverty status by marital status
#| echo: false
#| warning: false
#| message: false
poverty2019 |>
  ggplot(aes(x = marital_status, fill = poverty_status)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=6)) +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Marital status",
    y = "Number of respondents",
    fill = "Poverty status"
  )
```
```{r}
#| label: fig-income
#| fig-cap: The distribution of poverty status by income level
#| echo: false
#| warning: false
#| message: false
poverty2019 |>
  ggplot(aes(x = income, fill = poverty_status)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Income level",
    y = "Number of respondents",
    fill = "Poverty status"
  )
```

```{r}
#| label: fig-age
#| fig-cap: The distribution of poverty status by age group
#| echo: false
#| warning: false
#| message: false
poverty2019 |>
  ggplot(aes(x = age, fill = poverty_status)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Age group",
    y = "Number of respondents",
    fill = "Poverty status"
  )
```

@@

# Convert variables to factors
analysis_data$marital_status <- factor(analysis_data$marital_status)
analysis_data$income <- factor(analysis_data$income)
analysis_data$age <- factor(analysis_data$age)

@@

posterior_vs_prior(poverty_prediction_model, size = 0.1) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=5)) +
  guides(color = guide_legend(ncol = 4))+
  coord_flip()
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and marital status"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Married - civilian spouse present", "Trace plot of Married - armed forces spouse present", "Trace plot of Married - spouse absent (excluding separated)", "Trace plot of Widowed", "Trace plot of Separated", "Trace plot of Never Married"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "(Intercept)")
plot(poverty_prediction_model, "trace", "marital_statusMarried - armed forces spouse present")
plot(poverty_prediction_model, "trace", "marital_statusMarried - civilian spouse present")
plot(poverty_prediction_model, "trace", "marital_statusMarried - spouse absent (excluding separated)")
plot(poverty_prediction_model, "trace", "marital_statusNever Married")
plot(poverty_prediction_model, "trace", "marital_statusSeparated")
plot(poverty_prediction_model, "trace", "marital_statusWidowed")
```

@@

plot(poverty_prediction_model, "rhat") + xlim(-100,1.1)
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"
